Pre-Analysis Plan: Who Bends and Who Breaks? Experimental Evidence for Conditions Affecting Voters' Evaluations of Politicians Striking Compromises

### 2021-04-21 docs/pap/compromise-punishment-pap.Rmd
add updated pap

---
title: "Pre-Analysis Plan: Who Bends and Who Breaks? Experimental Evidence for Conditions Affecting Voters' Evaluations of Politicians Striking Compromises"
author: '16 April 2021'
output: 
  pdf_document:
    toc: yes
    number_sections: true
editor_options: 
  chunk_output_type: console
bibliography: references.bib
urlcolor: blue
linkcolor: blue
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage{xcolor}
---
```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
#library(printr)
library(tidyverse)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
```

@@

```{r power-example, eval = FALSE}
source(here::here("docs/pap/poweranalysis.R"))
```

@@

# Analysis

## Balance Checks
I will conduct a balance test based on demographics (age, gender, education, geographical region, level of urbanness, migration background,  employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political efficacy, attitudes towards women in politics, and positions on the four issues, using the `cobalt` R package [@greifer2021]. 
If the groups are unbalanced on one of these variables -- i.e. standardized mean differences `<0.05` -- I will add the covariates to the analyses.

@@

## Hypothesis 1
I test the hypotheses formulated in [Section 1](#expectations) applying a multiverse approach [@steegen2016increasing;@simonsohn2019specification] using the `specr` R package [@masur2020specr].
Instead of running various models using the different dependent variables

@@

@@

>>>>

---
title: "Pre-Analysis Plan: Who Bends and Who Breaks? Experimental Evidence for Conditions Affecting Voters' Evaluations of Politicians Striking Compromises"
date: "`r format(Sys.time(), '%d %B, %Y - %X (%Z)')`"
output: 
  pdf_document:
    toc: yes
    number_sections: true
editor_options: 
  chunk_output_type: console
bibliography: references.bib
urlcolor: blue
linkcolor: blue
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage{xcolor}
---
```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
#library(printr)
library(tidyverse)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
```

@@

```{r power-example, eval = FALSE, echo = TRUE}
source(here::here("docs/pap/poweranalysis.R"))
```

@@

# Analysis
I test the hypotheses formulated in [Section 1](#expectations) by fitting linear multivariate regressions separately for the four issues.
In each model, will estimate the coefficient for gender of the politician (H1), migration background of the politician (H2), and the interaction of gendere and migration background (H3).
I will apply a multiverse approach [@steegen2016increasing;@simonsohn2019specification].
Instead of running various models using the different dependent variables, I will estimate whether the treatment explains the different operationalizations of the dependent variable ([see Section 3.1](#dependent-variables)) in one go and present the estimates and confidence intervals for the several combinations.
I will only add control variables in the analyses that are unbalanced, as explained in [Section 4.1](#balance-checks).

## Balance Checks
I will conduct a balance test based on demographics (age, gender, education, geographical region, level of urbanness, migration background,  employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political efficacy, attitudes towards women in politics, and positions on the four issues, using the `cobalt` R package [@greifer2021]. 
If the groups are unbalanced on one of these variables -- i.e. standardized mean differences `<0.05` -- I will add the covariates to the analyses.
I will use the code below to conduct the balance tests (see [here](https://github.com/MarikenvdVelden/compromise-punish/tree/main/src/analysis/balance-test.R) for the R script).

@@

## Hypothesis 1
I test the _gender hypothesis_ using Equation \ref{eq:h1} and the code chunk below.
The coefficient of $\beta_{1}$ denotes the difference in the evaluation of female and male politicians.
The coefficient of $\beta_{2}$ denotes the difference in the evaluation of politicians who do and do not strike a comprmise.
The coefficient of $\beta_{3}$ is the interaction of the  $\beta_{1}$ and  $\beta_{2}$.
If the Average Marginal Effect of $\beta_{1}$ is negative and statistically siginificant when $\beta_{2} = 1$ (i.e. politician stroke a compromise), the _gender hypothesis_ will be confirmed.

\begin{equation}\label{eq:h1}
  \hat{Y} = \beta_{0} + \beta_{1}Female  + \beta_{2}Compromise + \beta_{3}Compromise * Female + \varepsilon
\end{equation}

```{r h1, eval = FALSE, echo = TRUE}
df <- d %>% select(PT1:PT4, gender, compromise)

depVarList <- setdiff(colnames(df), c("gender", "compromise"))
allModels <- lapply(depVarList, function(x){
  lm(formula= paste0("`", x, "` ~ gender * compromise"), 
     data= d, na.action = na.omit)})

for(i in 1:length(depVarList)){
  if(i==1){
    m <- summary(margins(allModels[[i]]), at = list(compromise = 1)) %>%
      filter(factor=="gender") %>%
      select(AME, lower, upper) %>%
      mutate(y = depVarList[i])
  }
  else{
    tmp <- summary(margins(allModels[[i]]), at = list(compromise = 1)) %>%
      filter(factor=="gender") %>%
      select(AME, lower, upper) %>%
      mutate(y = depVarList[i])
    m <- m %>%
      add_case(tmp)
  }
}
```

@@

I test the _migration hypothesis_ using Equation \ref{eq:h2} and the code chunk below.
The coefficient of $\beta_{1}$ denotes the difference in the evaluation of politicians with and without a migration background.
If $\beta_{1}$ is negative and statistically siginificant, the _migration hypothesis_ will be confirmed.
The coefficient of $\beta_{2}$ denotes the difference in the evaluation of politicians who do and do not strike a comprmise.
The coefficient of $\beta_{3}$ is the interaction of the  $\beta_{1}$ and  $\beta_{2}$.
If the Average Marginal Effect of $\beta_{1}$ is negative and statistically siginificant when $\beta_{2} = 1$ (i.e. politician stroke a compromise), the _migration hypothesis_ will be confirmed.

\begin{equation}\label{eq:h2}
  \hat{Y} = \beta_{0} + \beta_{1}Migration Background + \beta_{2}Compromise + \beta_{3}Compromise * Migration Background + \varepsilon
\end{equation}

```{r h2, eval = FALSE, echo = TRUE}
df <- d %>%
  mutate(migration = recode(name, 
                         `Rachid Amezian` = 1,
                         `Rachida Amezian` = 1,
                         .default = 0)) %>%
  select(PT1:PT4, migration, compromise)

depVarList <- setdiff(colnames(df), c("migration", "compromise"))
allModels <- lapply(depVarList, function(x){
  lm(formula= paste0("`", x, "` ~ migration * compromise"), 
     data= d, na.action = na.omit)})

for(i in 1:length(depVarList)){
  if(i==1){
    m <- summary(margins(allModels[[i]]), at = list(compromise = 1)) %>%
      filter(factor=="migration") %>%
      select(AME, lower, upper) %>%
      mutate(y = depVarList[i])
  }
  else{
    tmp <- summary(margins(allModels[[i]]), at = list(compromise = 1)) %>%
      filter(factor=="migration") %>%
      select(AME, lower, upper) %>%
      mutate(y = depVarList[i])
    m <- m %>%
      add_case(tmp)
  }
}
```

@@

I test the _intersection hypothesis_ using Equation \ref{eq:h3} and the code chunk below.
The coefficient of $\beta_{1}$ denotes the difference in the evaluation of female politicians with a migration background and the other politicians.
If $\beta_{1}$ is negative and statistically siginificant, the _gender hypothesis_ will be confirmed.
The coefficient of $\beta_{2}$ denotes the difference in the evaluation of politicians who do and do not strike a comprmise.
The coefficient of $\beta_{3}$ is the interaction of the  $\beta_{1}$ and  $\beta_{2}$.
If the Average Marginal Effect of $\beta_{1}$ is negative and statistically siginificant when $\beta_{2} = 1$ (i.e. politician stroke a compromise), the _intersection hypothesis_ will be confirmed.

\begin{equation}\label{eq:h3}
  \hat{Y} =\beta_{0} + \beta_{1}Intersection + \beta_{2}Compromise + \beta_{3}Compromise * Intersection + \varepsilon
\end{equation}

```{r h3, eval = FALSE, echo = TRUE}
df <- d %>%
  mutate(intersection = recode(name, 
                         `Rachida Amezian` = 1,
                         .default = 0)) %>%
  select(PT1:PT4, intersection, compromise)

depVarList <- setdiff(colnames(df), c("intersection", "compromise"))
allModels <- lapply(depVarList, function(x){
  lm(formula= paste0("`", x, "` ~ intersection * compromise"), 
     data= d, na.action = na.omit)})

for(i in 1:length(depVarList)){
  if(i==1){
    m <- summary(margins(allModels[[i]]), at = list(compromise = 1)) %>%
      filter(factor=="intersection") %>%
      select(AME, lower, upper) %>%
      mutate(y = depVarList[i])
  }
  else{
    tmp <- summary(margins(allModels[[i]]), at = list(compromise = 1)) %>%
      filter(factor=="intersection") %>%
      select(AME, lower, upper) %>%
      mutate(y = depVarList[i])
    m <- m %>%
      add_case(tmp)
  }
}
```

## Pooled Data
Using the pooled data we will estimate a within groups fixed effects model. 
$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled} denotes  the  evaluation  of  a politician of respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
$\alpha_{i}$ is the issue specific intercept, and $\gamma_{t}$ is the experimental round specific intercept.  
The standard errors are clustered at the individual level.  
If the Average Marginal Effect of $\beta_{1}$ is negative and statistically siginificant when $\beta_{2} = 1$ (i.e. politician stroke a compromise), the _intersection hypothesis_ will be confirmed.

\begin{equation}\label{eq:pooled}
    \hat{Y}_{r, i, t} =\beta_{0} + \beta_{1}Intersection_{r, i, t} + \beta_{2}Compromise + \beta_{3}Compromise * Intersection + \alpha_{i} + \gamma_{t} + \varepsilon_{r, i, t}
\end{equation}

## Statistical Significance
All the hypotheses are directional, and therefore all of the tests will be one-tailed.
I will use an $\alpha$-value of `0.05` as the value for statistical significance in all models above.
